### 2022-03-02 Manuscript/PARManuscript.tex
Add files via upload

% Created by Bonita Graham
% Last update: December 2019 By Kestutis Bendinskas
% Authors: 
% Please do not make changes to the preamble until after the solid line of %s.
\documentclass[10pt]{article}
\usepackage[explicit]{titlesec}
\setlength{\parindent}{0pt}
\setlength{\parskip}{1em}
\usepackage{hyphenat}
\usepackage{ragged2e}
\RaggedRight
% These commands change the font. If you do not have Garamond on your computer, you will need to install it.
\usepackage{garamondx}
\usepackage[T1]{fontenc}
\usepackage{amsmath, amsthm}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{xcolor}
\newcommand{\jy}[1]{\textcolor{orange}{JY: (#1)}}
\usepackage{booktabs}
% This adjusts the underline to be in keeping with word processors.
\usepackage{soul}
\setul{.6pt}{.4pt}
% The following sets margins to 1 in. on top and bottom and .75 in on left and right, and remove page numbers.
\usepackage{geometry}
\geometry{vmargin={1in,1in}, hmargin={.75in, .75in}}
\usepackage{fancyhdr}
\pagestyle{fancy}
\pagenumbering{gobble}
\renewcommand{\headrulewidth}{0.0pt}
\renewcommand{\footrulewidth}{0.0pt}
% These Commands create the label style for tables, figures and equations.
\usepackage[labelfont={footnotesize,bf} , textfont=footnotesize]{caption}
\captionsetup{labelformat=simple, labelsep=period}
\newcommand\num{\addtocounter{equation}{1}\tag{\theequation}}
\renewcommand{\theequation}{\arabic{equation}}
\makeatletter
\renewcommand\tagform@[1]{\maketag@@@ {\ignorespaces {\footnotesize{\textbf{Equation}}} #1.\unskip \@@italiccorr }}
\makeatother
\setlength{\intextsep}{10pt}
\setlength{\abovecaptionskip}{2pt}
\setlength{\belowcaptionskip}{-10pt}
\renewcommand{\textfraction}{0.10}
\renewcommand{\topfraction}{0.85}
\renewcommand{\bottomfraction}{0.85}
\renewcommand{\floatpagefraction}{0.90}
% These commands set the paragraph and line spacing
\titleformat{\section}
  {\normalfont}{\thesection}{1em}{\MakeUppercase{\textbf{#1}}}
\titlespacing\section{0pt}{0pt}{-10pt}
\titleformat{\subsection}
  {\normalfont}{\thesubsection}{1em}{\textit{#1}}
\titlespacing\subsection{0pt}{0pt}{-8pt}
\renewcommand{\baselinestretch}{1.15}
% This designs the title display style for the maketitle command
\makeatletter
\newcommand\sixteen{\@setfontsize\sixteen{17pt}{6}}
\renewcommand{\maketitle}{\bgroup\setlength{\parindent}{0pt}
\begin{flushleft}
\sixteen\bfseries \@title
\medskip
\end{flushleft}
\textit{\@author}
\egroup}
\makeatother
% This styles the bibliography and citations.
%\usepackage[biblabel]{cite}
\usepackage[sort&compress]{natbib}
\setlength\bibindent{2em}
\makeatletter
\renewcommand\@biblabel[1]{\textbf{#1.}\hfill}
\makeatother
\renewcommand{\citenumfont}[1]{\textbf{#1}}
\bibpunct{}{}{,~}{s}{,}{,}
\setlength{\bibsep}{0pt plus 0.3ex}
\usepackage{float}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Authors: Add additional packages and new commands here.  
% Limit your use of new commands and special formatting.
% Place your title below. Use Title Capitalization.
\title{Points Above Replacement: A New NBA Metric to Evaluate Player Performance}
% Add author information below. Communicating author is indicated by an asterisk, the affiliation is shown by superscripted lower case letter if several affiliations need to be noted.
\author{
Brian Krikorian*, Jun Yan \\ \medskip 
Department of Statistics, University of Connecticut, Storrs, CT  \\  \medskip 
Students: brian.krikorian@uconn.edu* \\
Mentor: jun.yan@uconn.edu 
}
\pagestyle{empty}
\begin{document}
% Makes the title and author information appear.
\vspace*{.01 in}
\maketitle
\vspace{.12 in}
% Abstracts are required.
\section*{abstract}
In attempting to predict the score of an NBA game, players who are injured, suspended, or for any other reason not playing have a significant impact on said prediction. Baseball uses a statistic called Wins Above Replacement, or WAR, to tell how many wins a team would gain or lose if they replaced a certain player with one who is league average. The goal of this paper is to create a similar statistic, called Points Above Replacement, or PAR, that uses statistics from throughout the season to predict how many points a team gains or loses by missing a certain player, and thus having to replace him with one who is "league average". The statistic is standardized and adjusted by both position and value to a player's given team. The model roughly follows the calculation of WAR in baseball, assigning values to points created on offense and defense, as well as a positional and team adjustment. A comparison with the created metric along with other well-known advanced player metrics is done via a multiple regression.
% Keywords are required.
\section*{keywords} 
NBA; Basketball; Player Performance; Basketball Metrics;
\vspace{.12 in}
% Start the main part of the manuscript here.
% Comment out section headings if inappropriate to your discipline.
% If you add additional section or subsection headings, use an asterisk * to avoid numbering. 
\section{Introduction}
There exists many prediction models for NBA basketball, many with the purpose of anticipating the final score of a given game. However, the majority of these models use team average statistics to base their decision, similar to the openWAR method in baseball introduced by Greg Matthews and others. However, teams rosters can change every night, and do often. When star players are ruled out, players of lower quality take their minutes, and are expected to perform at a lower level. When lower level players do not play in a given game, star players are expected to play more minutes, which most likely will cause a slight uptick in performance. This reasoning is the cause for the creation of the new metric, PAR. The measure of how many points a team will gain or lose if a certain player does not play. The metric is created based on the last completed season, the 2020-2021 season, which was cut to 72 games, as opposed to 82, due to the COVID-19 pandemic. Future models could be adjusted to a full 82 game season, and automated to update statistics as the season goes on. Position and team adjustments allow for a player's value to his specific team to be measured, giving a better expectation for their presence or absence. Adjusting to a position allows one to analyze a player based on what his position requires. A point guard is expected to pass, shoot three pointers, and get steals more than he is expected to block shots and rebound. The opposite is true for centers, so different coefficients are created based on the primary position a player played throughout the season. The team adjustment is a function of how many minutes a player played throughout the entire season, and how many of their team's total points were scored by them. If a player plays a majority of his team's minutes, and scores a relatively high percentage of his team's points, he will be more valuable to his team, and his team will be hurt more by his absence in a game.
\section{Literature Review}
\section{Data}
All basic and advanced statistics from the 2020-2021 season were collected from the Basketball Reference website. Each teams total points throughout the course of the season were also collected from the Basketball Reference website. From the basic statistics, 4 metrics were created in order to calculate PAR. Offensive Points is a function of how many points a player will create on offense. While the coefficients vary by position, the statistics used are effective field goal percentage, three point percentage, free throw percentage, offensive rebounds, assists, points, and turnovers. Similarly, Defensive Points is a function of how many points a player helps prevent while his team is on defense. The coefficients also vary by position, but this metric is based on defensive rebounds, total rebounds, steals, blocks, and personal fouls. The Positional Adjustment metric creates a coefficient of how valuable a player is to their individual team, and is based on how many minutes a player plays relative to their team, as well as how many points they score relative to their teams total points. The last statistic created is Points Per Win, which is based on the average amount of points a player will create in a game. This function is based on the sum of Offensive and Defensive Points, and is scaled to incorporate the total number of games, minutes, and players in a season. This statistic is very similarly calculated to Runs Per Win in the WAR statistic. Given that over the course of a 72 game season, all teams will play each other at least once, we assume a relatively equal distribution in terms of strength of schedule. Advanced statistics for each player were also collected in order to perform a regression, to see how PAR compares with known and accepted metrics. Value Over Replacement Player, also known as VORP, was one of the metrics used, and is calculated with the help of Box Score Plus Minus, also known as BPM. These statistics also help give a relative estimate of a player's value compared to that of an average player, but they translate the statistic to a league average team as well. PAR does not do this, because it is concerned with how valuable a player is to his specific team, not throughout the entire league. If a player switched teams, his PAR value would most likely change, as his importance to one team may be more or less than it would to another. For simplicity, the team that a player played his most games on was taken into consideration for this metric. Other statistics used include Player Efficiency Rating, a per-minute rating of player performance, and Win Shares, which estimates the number of wins a player creates for his team throughout the season. Due to all of these metrics' relative similarity to PAR, it is suggested that if they are highly correlated with PAR, it is a successful metric. However, we expect some differentiation between the metrics, as they all vary slightly in what they attempt to convey.
\section{Methods}
\section{Results}
\section{Discussion}
\bibliographystyle{chicago}
\bibliography{citations.bib}
% \section*{acknowledgements}
% The authors thank the University of Connecticut and the UConn Department of Statistics.
%Note correct LaTeX quotations above. Do not use the " symbol, but rather double ` followed by double '
% The About the Student Author section is NOT optional.  Write a paragraph about the student; see previous journal editions for examples.
% If there is more than one student author, you must move the comment below.
\section*{about the student author}
%\section*{about the student authors}
% The Press Summary section is NOT optional.  Write a paragraph describing the paper in a manner suitable for the press; see previous journal editions for examples.
\section*{press summary}
\end{document}