### 2022-03-21 Manuscript/PARManuscript.tex
Add files via upload

% Created by Bonita Graham
% Last update: December 2019 By Kestutis Bendinskas
% Authors: 
% Please do not make changes to the preamble until after the solid line of %s.
\documentclass[10pt]{article}
\usepackage[explicit]{titlesec}
\setlength{\parindent}{0pt}
\setlength{\parskip}{1em}
\usepackage{hyphenat}
\usepackage{ragged2e}
\RaggedRight
% These commands change the font. If you do not have Garamond on your computer, you will need to install it.
\usepackage{garamondx}
\usepackage[T1]{fontenc}
\usepackage{amsmath, amsthm}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{xcolor}
\newcommand{\jy}[1]{\textcolor{orange}{JY: (#1)}}
\usepackage{booktabs}
% This adjusts the underline to be in keeping with word processors.
\usepackage{soul}
\setul{.6pt}{.4pt}
% The following sets margins to 1 in. on top and bottom and .75 in on left and right, and remove page numbers.
\usepackage{geometry}
\geometry{vmargin={1in,1in}, hmargin={.75in, .75in}}
\usepackage{fancyhdr}
\pagestyle{fancy}
\pagenumbering{gobble}
\renewcommand{\headrulewidth}{0.0pt}
\renewcommand{\footrulewidth}{0.0pt}
% These Commands create the label style for tables, figures and equations.
\usepackage[labelfont={footnotesize,bf} , textfont=footnotesize]{caption}
\captionsetup{labelformat=simple, labelsep=period}
\newcommand\num{\addtocounter{equation}{1}\tag{\theequation}}
\renewcommand{\theequation}{\arabic{equation}}
\makeatletter
\renewcommand\tagform@[1]{\maketag@@@ {\ignorespaces {\footnotesize{\textbf{Equation}}} #1.\unskip \@@italiccorr }}
\makeatother
\setlength{\intextsep}{10pt}
\setlength{\abovecaptionskip}{2pt}
\setlength{\belowcaptionskip}{-10pt}
\renewcommand{\textfraction}{0.10}
\renewcommand{\topfraction}{0.85}
\renewcommand{\bottomfraction}{0.85}
\renewcommand{\floatpagefraction}{0.90}
% These commands set the paragraph and line spacing
\titleformat{\section}
  {\normalfont}{\thesection}{1em}{\MakeUppercase{\textbf{#1}}}
\titlespacing\section{0pt}{0pt}{-10pt}
\titleformat{\subsection}
  {\normalfont}{\thesubsection}{1em}{\textit{#1}}
\titlespacing\subsection{0pt}{0pt}{-8pt}
\renewcommand{\baselinestretch}{1.15}
% This designs the title display style for the maketitle command
\makeatletter
\newcommand\sixteen{\@setfontsize\sixteen{17pt}{6}}
\renewcommand{\maketitle}{\bgroup\setlength{\parindent}{0pt}
\begin{flushleft}
\sixteen\bfseries \@title
\medskip
\end{flushleft}
\textit{\@author}
\egroup}
\makeatother
% This styles the bibliography and citations.
%\usepackage[biblabel]{cite}
\usepackage[sort&compress]{natbib}
\setlength\bibindent{2em}
\makeatletter
\renewcommand\@biblabel[1]{\textbf{#1.}\hfill}
\makeatother
\renewcommand{\citenumfont}[1]{\textbf{#1}}
\bibpunct{}{}{,~}{s}{,}{,}
\setlength{\bibsep}{0pt plus 0.3ex}
\usepackage{float}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Authors: Add additional packages and new commands here.  
% Limit your use of new commands and special formatting.
% Place your title below. Use Title Capitalization.
\title{Points Above Replacement: A New NBA Metric to Evaluate Player Performance}
% Add author information below. Communicating author is indicated by an asterisk, the affiliation is shown by superscripted lower case letter if several affiliations need to be noted.
\author{
Brian Krikorian*, Jun Yan \\ \medskip 
Department of Statistics, University of Connecticut, Storrs, CT  \\  \medskip 
Students: brian.krikorian@uconn.edu* \\
Mentor: jun.yan@uconn.edu 
}
\pagestyle{empty}
\begin{document}
% Makes the title and author information appear.
\vspace*{.01 in}
\maketitle
\vspace{.12 in}
% Abstracts are required.
\section*{abstract}
In attempting to predict the score of an NBA game, players who are injured, suspended, or for any other reason not playing have a significant impact on said prediction. Baseball uses a statistic called Wins Above Replacement, or WAR, to tell how many wins a 
team would gain or lose if they replaced a certain player with one who is league average. The goal of this paper is to create a similar statistic, called Points Above Replacement, or PAR, that uses statistics from throughout the season to predict how many points a 
team gains or loses by missing a certain player, and thus having to replace him with one who is "league average". The statistic is standardized and adjusted by both position and value to a player's given team. The model roughly follows the calculation of WAR in 
baseball, assigning values to points created on offense and defense, as well as a positional and team adjustment. A comparison with the created metric along with other well-known advanced player metrics is done via a multiple regression.
% Keywords are required.
\section*{keywords} 
NBA; Basketball; Player Performance; Basketball Metrics;
\vspace{.12 in}
% Start the main part of the manuscript here.
% Comment out section headings if inappropriate to your discipline.
% If you add additional section or subsection headings, use an asterisk * to avoid numbering.

\section{Introduction}
%Why is this important/interesting? What has been done before on this topic? (Literature Review) What am I introducing that is new? 
There exists many prediction models for NBA basketball, many with the purpose of anticipating the final score of a given game. However, the majority of these models use team average statistics to base their decision, similar to the openWAR method in baseball 
introduced by Greg Matthews and others. However, teams rosters can change every night, and do often. When star players are ruled out, players of lower quality take their minutes, and are expected to perform at a lower level. When lower level players do not 
play in a given game, star players are expected to play more minutes, which most likely will cause a slight uptick in performance. This reasoning is the cause for the creation of the new metric, PAR. The measure of how many points a team will gain or lose if a 
certain player does not play. The metric is created based on the last completed season, the 2020-2021 season, which was cut to 72 games, as opposed to 82, due to the COVID-19 pandemic. Future models could be adjusted to a full 82 game season, and 
automated to update statistics as the season goes on. Position and team adjustments allow for a player's value to his specific team to be measured, giving a better expectation for their presence or absence. Adjusting to a position allows one to analyze a player 
based on what his position requires. A point guard is expected to pass, shoot three pointers, and get steals more than he is expected to block shots and rebound. The opposite is true for centers, so different coefficients are created based on the primary position 
a player played throughout the season. The team adjustment is a function of how many minutes a player played throughout the entire season, and how many of their team's total points were scored by them. If a player plays a majority of his team's minutes, and 
scores a relatively high percentage of his team's points, he will be more valuable to his team, and his team will be hurt more by his absence in a game.

\section{Data}
All basic and advanced statistics from the 2020-2021 season were collected from the Basketball Reference website. Each teams total points throughout the course of the season were also collected from the Basketball Reference website.

From the basic statistics, 4 metrics were created in order to calculate PAR. Offensive Points is a function of how many points a player will create on offense. While the coefficients vary by position, the statistics used are effective field goal percentage, three point 
percentage, free throw percentage, offensive rebounds, assists, points, and turnovers. Similarly, Defensive Points is a function of how many points a player helps prevent while his team is on defense. The coefficients also vary by position, but this metric is based 
on defensive rebounds, total rebounds, steals, blocks, and personal fouls. The Positional Adjustment metric creates a coefficient of how valuable a player is to their individual team, and is based on how many minutes a player plays relative to their team, as well 
as how many points they score relative to their teams total points. The last statistic created is Points Per Win, which is based on the average amount of points a player will create in a game. This function is based on the sum of Offensive and Defensive Points, 
and is scaled to incorporate the total number of games, minutes, and players in a season. This statistic is very similarly calculated to Runs Per Win in the WAR statistic.

Given that over the course of a 72 game season, all teams will play each other at least once, we assume a relatively equal distribution in terms of strength of schedule. Advanced statistics for each player were also collected in order to perform a regression, to 
see how PAR compares with known and accepted metrics. Value Over Replacement Player, also known as VORP, was one of the metrics used, and is calculated with the help of Box Score Plus Minus, also known as BPM. These statistics also help give a 
relative estimate of a player's value compared to that of an average player, but they translate the statistic to a league average team as well. PAR does not do this, because it is concerned with how valuable a player is to his specific team, not throughout the entire 
league. If a player switched teams, his PAR value would most likely change, as his importance to one team may be more or less than it would to another. For simplicity, the team that a player played his most games on was taken into consideration for this metric. 
Other statistics used include Player Efficiency Rating, a per-minute rating of player performance, and Win Shares, which estimates the number of wins a player creates for his team throughout the season. Due to all of these metrics' relative similarity to PAR, it is 
suggested that if they are highly correlated with PAR, it is a successful metric. However, we expect some differentiation between the metrics, as they all vary slightly in what they attempt to convey. The relative correlation between all these statistics shows that 
PAR is related to other advanced metrics analyzing player efficiency, but also that it introduces something new in basketball statistics.

\section{Methods}
The basic method for calculating the PAR method was based on how WAR is calculated in baseball, but also took influences from other advanced basketball statistics. WAR is based on Offensive and Defensive Runs, which are calculated subjectively via 
attributes of the game such as hitting, base running, fielding, and so forth. The subjectivity is in the coefficients chosen to scale statistics which impact Offensive and Defensive Runs. Thus, there are actually an infinite different kinds of WAR in baseball, 
depending on which method of calculating Offensive and Defensive Runs is used. Similarly, PAR is created by calculating Offensive and Defensive Points statistics, to analyze how many points a player adds for his team, and how many he prevents the other 
team from scoring. Using the basic offensive statistics of Effective Field Goal Percentage, Three Point Percentage, Free Throw Percentage, Offensive Rebounds, Assists, Points, and Turnovers, we assign a coefficient to them based on how important they are for 
a given position. Point guards are expected to pass more than centers, and a point guard's passing ability has more value to his team than a center, so more weight is placed on that statistic for point guards. This goes for all positions, and a coefficient matrix of a 
constant scale is created. That is to say, assists have a coefficient from point guards, shooting guards, small forwards, power forwards, and centers, respectively, of 4, 3.5, 3, 2.5, and 2. The coefficient by position method is one similar to that of Box Plus Minus, 
another advanced NBA metric. The position a player was assigned for the calculation was the position where he played the most minutes throughout the 2020-2021 season. This coefficient method gives a clean scaling into the calculation of Offensive Points. 
Some statistics, such as Points and Effective Field Goal Percentage, are equally important, regardless of the position of a given player. Thus, these statistics have a constant coefficient that does not change. Once all coefficients were applied to all statistics, the 
sum was taken to give Offensive Points. The exact same method was applied to the Defensive Points calculation, but used the basic defensive statistics of Defensive Rebounds, Total Rebounds, Steals, Blocks, and Personal Fouls. As turnovers and personal 
fouls are not desirable amongst players, the coefficients for these statistics is negative, which allows for a scale of not only the good a player does, but also what they do poorly. This coefficient method is similar to the Positional Adjustment metric applied in WAR 
calculation, which seeks to adjust a player's importance based on what is expected of their position.

The next step was creating a Team Adjustment statistic. As stated before, the PAR metric seeks to anticipate how valuable a player is to their given team, rather than as an average of the entire league. This gives more of an insight of how a team will fare without 
a certain player. We expect this statistic to change if a player were to switch teams. This is similar to the League Adjustment in WAR calculation, where a player's value is adjusted based on whether he plays in the American League or the National League. The 
Eastern and Western Conferences in the NBA have much less disparity compared to the MLB, and each conference plays each other much more frequently, so we instead change this to adjust a player's value to his specific team. The team assigned to a given 
player was the team where he played the most games during the 2020-2021 season. If a player had played an even amount of games for 2 separate teams, the team the player had played the most minutes for was assigned as his team. The Team Adjustment 
statistic is then calculated by taking the total minutes a player was on the court during the entire season, and diving it by a constant of 3,456 (the maximum number of minutes a player could have played during the shortened season, 48 minutes per game 
multiplied by 72 games). Then, the percentage of points a player scored for his team (player's total points divided by team's total points throughout the season) was calculated, and the two results were multiplied together. As a player cannot play 100 percent of 
his team's minutes, or score 100 percent of his team's points, the result is a number between 0 and 1. A similar adjustment is also used in the calculation of Box Plus Minus, which is a key factor in calculating Value Over Replacement Player, two of the advanced 
metrics PAR will be compared against. The goal of this adjustment is to give less of a penalty to players who play a majority of their team's minutes, and score a majority of their team's points. This helps truly elite players stand out amongst the pack, as well as 
players who may not be the best of the best, but matter more to their team specifically than in the grand scheme of things.

Lastly, WAR implements a Runs Per Win statistic, which gives a basic idea of how many runs are needed for a win in the MLB. Similarly, a Points Per Win statistic is created for the NBA, which allows a scaling down of the PAR statistic. We calculate this via the 
combined Offensive and Defensive Points statistic for a player divided by the number of minutes he played, then adding 2,160 (72 games per season times 30 teams in the NBA, thus, the total number of chances to win throughout the season). This gives a basic 
metric of a players points created as a function of how many minutes he played, and allows us to even the metric out a bit more from star players to role players. This statistic serves as the denominator of our PAR calculation.

Once all supporting statistics of PAR have been created, all that is left to do is combine them to achieve our PAR values for each player in the NBA. The formula for PAR is as follows:

\begin{equation}
((Offensive Points + DefensivePoints) * (Team Adjustment))/(Points Per Win)
\end{equation}

This gives a good idea of how many points a player accounts for during a game, how important that is for his given team, and how important it is compared to all other players. The last step is to standardize this PAR statistic, so that the mean is 0, and the 
standard deviation is 1. Logically, this helps the idea of the metric in the sense that a perfectly league average player should have a PAR of 0, and his presence or absence from his team neither adds nor subtracts value. Thus, we have created a new NBA player 
metric, Points Above Replacement, PAR.

\section{Results}
% Multiple Regression
% Analyze Graphs of PAR, filter by position perhaps?
% Any other tests that should be done?

\section{Discussion}

\bibliographystyle{chicago}
%\bibliography{citations.bib}
% \section*{acknowledgements}
The authors thank the University of Connecticut and the UConn Department of Statistics.
%Note correct LaTeX quotations above. Do not use the " symbol, but rather double ` followed by double '
% The About the Student Author section is NOT optional.  Write a paragraph about the student; see previous journal editions for examples.
% If there is more than one student author, you must move the comment below.
\section*{about the student author}
%\section*{about the student authors}
Brian Krikorian plans to graduate in May 2022 from the University of Connecticut, with a Bachelor's of Science in Data Science, and a minor in Computer Science.
% The Press Summary section is NOT optional.  Write a paragraph describing the paper in a manner suitable for the press; see previous journal editions for examples.
\section*{press summary}
\end{document}